---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file, then run
     devtools::build_readme() -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# classir

Note that this package was entirely coded by Claude AI based on back-and-forth
exchanges with me. Much of the text in the documentation was also written by
Claude AI. The package is also very much a work in progress and hasn't been
thoroughly checked yet. It's worth double checking all its output.

This is a package for the typical kinds of operations psychologists (and other
researchers) do in R. It has functions for surveying datasets (`look()`) and
combining datasets (`add_values()`). It also has some functions that are more
geared towards psycholinguistic analyses but might be helpful for others. This
includes a function to simplify a maximally complex random effects structure of
a linear mixed effects model (`converge()`) and a function that runs multiple
approaches to building regression models and reports all of the results
(`compare_methods()`). It also has a function to transcribe words into IPA
(`transcribe_phonemes()`).

I built this to streamline the operations I did often in R. I hope it will be
useful for you too!

The package is named after my lab, the Cognition, Language, Sound Symbolism,
Iconicity (CLaSSI) Lab. If this is helpful for you I'd love to hear about it
and can be reached at david.sidhu@carleton.ca

## Installation

```r
install.packages("remotes")
remotes::install_github("davidsidhu/classir")
```

If R asks whether you want to install a package from source, answer **no** —
saying yes will try to compile it, which usually fails without extra tools.

Some functions need packages that aren't installed automatically:

```r
install.packages(c("afex", "lmerTest", "VSURF", "readxl"))
```

## Start here

The four functions that will be helpful to most people are:

- **`look()`** — a survey of your data, one variable at a time
- **`check_data()`** — looks for anomalous cells
- **`prep()`** — prepares data for analyses
- **`add_values()`** — adds variables to your dataset, similar to VLOOKUP in
  Excel

The examples below use `demo_words`, a simulated dataset of 500 invented words
that comes with the package.

### `look()`

```{r}
library(classir)

look(demo_words)
```

### `check_data()`

```{r}
check_data(demo_words, id = "word")
```

### `add_values()`

```{r}
stimuli <- data.frame(word = demo_words$word[1:6])

add_values(stimuli, "word", demo_words, "word", c("freq", "aoa"))
```

Students who would rather answer questions than write the call can use
`add_values_guided()`, which walks through the same steps and prints the
equivalent command at the end.

## All functions

### Surveying and cleaning

| Function | What it does |
|---|---|
| `look()` | A short labelled block for each variable in your dataset: mean, SD, range, level counts, missingness |
| `check_data()` | Scans for likely problems — wrong types, heavy missingness, constant columns, strange cells |
| `prep()` | Prepares a dataset for analysis: drops incomplete rows, keeps a subset of rows, sets reference levels, effects codes two-level factors, standardises numeric variables and renames columns |
| `prep_guided()` | An interactive version of `prep()` |

### Combining datasets

| Function | What it does |
|---|---|
| `add_values()` | Adds columns from another dataset, matching rows on a shared column |
| `add_values_guided()` | An interactive version of `add_values()` |

### Describing

| Function | What it does |
|---|---|
| `by_group()` | Mean and SD table of selected variables by one or two grouping variables |

### Regression and variable selection

| Function | What it does |
|---|---|
| `compare_methods()` | Runs several selection methods (correlation, regression, stepwise regression, best subsets regression, adaptive lasso and random forests) on the same data and reports all of the results in a grid |
| `alasso()` | Adaptive lasso, with ridge-derived penalty weights |
| `best_subsets_boot()` | Best-subsets regression by BIC, with bootstrap selection frequencies |
| `best_subsets_perm()` | A permutation null for the above |
| `best_subsets_report()` | Prints a `best_subsets_boot()` result |
| `converge()` | Works through a sequence of steps to get a non-converging `lmer` model to converge |

### Phonemes

| Function | What it does |
|---|---|
| `transcribe_phonemes()` | Looks words up in a pronunciation dictionary and optionally codes each phoneme |
| `word_to_ipa()` | Spelled words to IPA |
| `code_phonemes()` | IPA strings to one column per phoneme |
| `cmu_dict()` | Downloads and caches the CMU pronouncing dictionary |
| `build_cmu_ipa()` | Builds a word-to-IPA table from a CMUdict file |
| `english_phonemes` | The CMUdict English phoneme inventory, in IPA |

### Data

| Object | What it is |
|---|---|
| `demo_words` | A simulated dataset of 500 invented words, used in these examples |

## Comparing methods

`compare_methods()` runs several variable selection approaches on the same
data and shows where they agree.

```{r}
compare_methods(demo_words, "iconicity",
                c("freq", "length", "old20", "aoa", "concreteness"),
                methods = c("cor", "lm", "step", "alasso", "subsets"),
                seed = 2026)
```

It is meant for examining how robust an effect is across analysis methods. It
is not for choosing which analysis to report based on the pattern of results.

## A note on how this was made

This package was entirely created by Claude AI based on direction from me.

## Credits

`cmu_dict()` uses the [CMU Pronouncing
Dictionary](https://github.com/cmusphinx/cmudict), Copyright (C) 1993–2014
Carnegie Mellon University.
